---
title: "p8105_mtp_Cr3442"
author: "Cheng Rao"
date: "2024-10-24"
output: github_document
---

```{r setup, include = FALSE}
library(tidyverse)
library(readxl)
library(janitor)
library(patchwork)
library(rvest)
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.width = 10,
  fig.height = 8,
  out.width = "90%"
)
theme_set(theme_minimal() + theme(legend.position = "bottom"))
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```



# Report

## Problem 1 – Data Import and Cleaning

### Introduction

My original data is the New York City Rental Price Index (ZORI) published by Zillow and the Zillow Home Price Index (ZHVI) for various regions, as well as information on different communities in different districts. I will analyze the intrinsic relationship between rents and house prices in some areas of New York and the corresponding years.

### Step 1

```{r import_clean_nyc_rental}
# Import NYC Rental Data and ZIP Code Data
nyc_rental = read_csv("./data/Zip_zori_uc_sfrcondomfr_sm_month_NYC.csv") %>%
  janitor::clean_names()

# Import Supplementary Dataset to Get Borough and Neighborhood Information
zip_data = read_html("./data/supplementary dataset.html") %>%
  html_table() %>%
  .[[1]] %>%
  janitor::clean_names() %>%
  mutate(
    borough = case_when(
      county == "New York" ~ "Manhattan",
      county == "Kings" ~ "Brooklyn",
      county == "Queens" ~ "Queens",
      county == "Bronx" ~ "Bronx",
      county == "Richmond" ~ "Staten Island"
    )
  )

# Standardize ZIP Code Column Name
nyc_rental = nyc_rental %>%
  rename(zip_code = region_name) %>%
  mutate(zip_code = as.character(zip_code))

# Ensure ZIP Code Column in Supplementary Data is Character Type
zip_data = zip_data %>%
  mutate(zip_code = as.character(zip_code))

# Merge NYC Rental Data with ZIP Code Data
nyc_data = nyc_rental %>%
  left_join(zip_data, by = "zip_code") %>%
  filter(!is.na(borough)) %>%  # Filter Out Rows Without Matching Borough
  select(zip_code, neighborhood, borough, county, starts_with("x")) 

# Count the total number of observations in the combined dataset
total_observations = nyc_data %>%
nrow()

# Count the number of unique zip codes
unique_zip_codes = nyc_data %>%
distinct(zip_code) %>%
nrow()

# Count the number of unique neighborhoods
unique_neighborhoods = nyc_data %>%
distinct(neighborhood) %>%
nrow()

# Print the results
cat("Total number of observations in the combined dataset:", total_observations, "\n")
cat("Number of unique zip codes in the combined dataset:", unique_zip_codes, "\n")
cat("Number of unique neighborhoods in the combined dataset:", unique_neighborhoods, "\n")
```


### Step 2

```{r import_clean_zhvi}
# Import ZHVI Data
zhvi_data = read_csv("./data/Zip_zhvi_uc_sfrcondo_tier_0.33_0.67_sm_sa_month_2023.csv") %>%
  janitor::clean_names() %>%
  rename(zip_code = region_name) %>%
  mutate(zip_code = as.character(zip_code))

# Convert ZHVI Data to Long Format
zhvi_long = zhvi_data %>%
  pivot_longer(
    cols = starts_with("x"),
    names_to = "date",
    values_to = "zhvi_price"
  ) %>%
  mutate(date = as.Date(date, format = "x%Y_%m_%d"))  
```

### Step 3

```{r pivot_zori}
# Convert ZORI Data to Long Format
zori_long = nyc_data %>%
  pivot_longer(
    cols = starts_with("x"),
    names_to = "date",
    values_to = "zori_rental_price"
  ) %>%
  mutate(date = as.Date(date, format = "x%Y_%m_%d")) 
```

Explanation: I first imported the three datasets, then cleaned and merged them based on standardized column names, then checked the merged dataset to ensure data consistency and integrity, and finally standardized the date column.

## Problem 2 – Quality Control and EDA

### Question 1

This may be because rental data may not have been recorded consistently for all ZIP codes over the entire period, or because ZIP codes with fewer rental properties may not have enough data to provide complete observations.

### Question 2

```{r compare_number}
# Count ZIP Codes in NYC Rental Dataset
num_zip_rental = nyc_rental %>%
  distinct(zip_code) %>%
  nrow()

# Count ZIP Codes in ZIP Code Dataset
num_zip_data = zip_data %>%
  distinct(zip_code) %>%
  nrow()

# Print the counts
cat("Number of ZIP Codes in NYC Rental Dataset:", num_zip_rental, "\n")
cat("Number of ZIP Codes in ZIP Code Dataset:", num_zip_data, "\n")
```

Comments: This could be because not all zip codes in New York City have enough rental properties to be included in the dataset, such as those with fewer rental listings.

### Question 3

```{r avg_rent_by_borough}
# Calculate Average Rental Price by Borough and Year
avg_rent_by_borough_year = zori_long %>%
  mutate(year = lubridate::year(date)) %>%
  group_by(borough, year) %>%
  summarise(avg_rental_price = mean(zori_rental_price, na.rm = TRUE)) %>%
  arrange(borough, year)

# Display Table
avg_rent_by_borough_year %>%
  knitr::kable(digits = 2, caption = "Average Rental Price by Borough and Year")
```

Comments: Overall, rental prices in all boroughs showed an upward trend between 2015 and 2024. From 2020 to 2021, rental prices declined, especially in Manhattan and Brooklyn, reflecting the lower demand in the rental market during the COVID-19 pandemic. Since 2022, rents have begun to recover significantly.

### Question 4

```{r rental_price_drop}
# Compare Rental Prices in January 2021 to January 2020
covid_rent_drop = zori_long %>%
  filter(date == as.Date("2020-01-31") | date == as.Date("2021-01-31")) %>%
  pivot_wider(names_from = date, values_from = zori_rental_price) %>%
  mutate(price_drop = `2020-01-31` - `2021-01-31`) %>%
  group_by(borough) %>%
  filter(price_drop == max(price_drop, na.rm = TRUE)) %>%
  select(borough, neighborhood, price_drop)

# Display Table
covid_rent_drop %>%
  knitr::kable(digits = 2, caption = "Largest Rental Price Drop from 2020 to 2021 by Borough")
```

Comments: According to the table, rental prices in Manhattan and Brooklyn were more significantly impacted during the epidemic, which may be because they are located in the downtown area . Relatively speaking, the Bronx and Queens were less affected, which may be because the environment there is more suitable for remote work.

## Problem 3 – Visualization

### Qusetion 1

```{r rental_price_plot}
# Plot NYC rental prices for each ZIP code over time
zori_long %>%
  ggplot(aes(x = date, y = zori_rental_price, color = borough)) +
  geom_line(alpha = 0.5) +
  labs(
    title = "NYC Rental Prices by ZIP Code Over Time",
    x = "Date",
    y = "Rental Price Index (ZORI)"
  ) +
  theme_minimal()
```

Comments: Overall, rental prices in all boroughs of New York City have been on an upward trend, but between 2020 and 2021, rental prices have dropped significantly in all boroughs. I think this is related to the trend of a large number of residents moving out of the city and choosing to work remotely during the epidemic. After that, rents quickly rebounded in 2022 and continued to rise.

### Qusetion 2

```{r zhvi_distribution_plot}
# Calculate average housing price for each ZIP code in 2023
avg_zhvi_by_zip = zhvi_long %>%
  group_by(zip_code) %>%
  summarise(avg_zhvi_price = mean(zhvi_price, na.rm = TRUE))

# Plot the distribution of housing prices to compare across states
avg_zhvi_by_zip %>%
  ggplot(aes(x = avg_zhvi_price)) +
  geom_histogram(binwidth = 50000, fill = "skyblue", color = "black", alpha = 0.7) +
  labs(
    title = "Distribution of ZIP Code-Level Housing Prices Across States in 2023",
    x = "Average Housing Price (ZHVI)",
    y = "Number of ZIP Codes"
  ) +
  theme_minimal()
```

Comments: The housing price curve shown in the figure shows a typical right-skewed distribution, with the average housing prices in most zip codes concentrated in the lower range, and only a few zip codes have very high housing prices.

### Qusetion 3

```{r rental_vs_housing_plot}
# Calculate the average monthly rental price and housing price for each ZIP code in 2023
zori_zhvi_2023 = zori_long %>%
  filter(lubridate::year(date) == 2023) %>%
  group_by(zip_code) %>%
  summarise(avg_rental_price = mean(zori_rental_price, na.rm = TRUE)) %>%
  left_join(
    zhvi_long %>%
      filter(lubridate::year(date) == 2023) %>%
      group_by(zip_code) %>%
      summarise(avg_zhvi_price = mean(zhvi_price, na.rm = TRUE)),
    by = "zip_code"
  )

# Plot rental prices against housing prices
zori_zhvi_2023 %>%
  ggplot(aes(x = avg_rental_price, y = avg_zhvi_price)) +
  geom_point(alpha = 0.6) +
  labs(
    title = "Comparison of Rental and Housing Prices by ZIP Code in NYC for 2023",
    x = "Average Rental Price (ZORI)",
    y = "Average Housing Price (ZHVI)"
  ) +
  theme_minimal()
```

Comments: From the scatter plot, we can see that there is a clear positive correlation between rental prices and housing prices. Most of the data points are concentrated in the lower range of rental and housing prices. Only a few data points correspond to higher rental and housing prices.

### Question 4

1. Rent and housing data may be incomplete for some zip codes, especially in areas with relatively low rental activity. 

2. The dataset is aggregated by zip code, which is likely to ignore price differences within zip codes. 


