---
title: "p8105_mtp_Cr3442"
author: "Cheng Rao"
date: "2024-10-24"
output: github_document
---

```{r setup, include = FALSE}
library(tidyverse)
library(readxl)
library(janitor)
library(patchwork)
library(rvest)
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.width = 10,
  fig.height = 8,
  out.width = "90%"
)
theme_set(theme_minimal() + theme(legend.position = "bottom"))
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```



# Report

## Problem 1 – Data Import and Cleaning

### Introduction

First, I imported and clean the data by 3 steps.

### Step 1: Import and Clean NYC Rental Data

```{r import_clean_nyc_rental}
# Import NYC Rental Data and ZIP Code Data
nyc_rental = read_csv("./data/Zip_zori_uc_sfrcondomfr_sm_month_NYC.csv") %>%
  janitor::clean_names()

# Import Supplementary Dataset to Get Borough and Neighborhood Information
zip_data = read_html("./data/supplementary dataset.html") %>%
  html_table() %>%
  .[[1]] %>%
  janitor::clean_names() %>%
  mutate(
    borough = case_when(
      county == "New York" ~ "Manhattan",
      county == "Kings" ~ "Brooklyn",
      county == "Queens" ~ "Queens",
      county == "Bronx" ~ "Bronx",
      county == "Richmond" ~ "Staten Island"
    )
  )

# Standardize ZIP Code Column Name
nyc_rental = nyc_rental %>%
  rename(zip_code = region_name) %>%
  mutate(zip_code = as.character(zip_code))

# Ensure ZIP Code Column in Supplementary Data is Character Type
zip_data = zip_data %>%
  mutate(zip_code = as.character(zip_code))

# Merge NYC Rental Data with ZIP Code Data
nyc_data = nyc_rental %>%
  left_join(zip_data, by = "zip_code") %>%
  filter(!is.na(borough)) %>%  # Filter Out Rows Without Matching Borough
  select(zip_code, neighborhood, borough, county, starts_with("x"))  
```

### Step 2: Import and Clean ZHVI Data

```{r import_clean_zhvi}
# Import ZHVI Data
zhvi_data = read_csv("./data/Zip_zhvi_uc_sfrcondo_tier_0.33_0.67_sm_sa_month_2023.csv") %>%
  janitor::clean_names() %>%
  rename(zip_code = region_name) %>%
  mutate(zip_code = as.character(zip_code))

# Convert ZHVI Data to Long Format
zhvi_long = zhvi_data %>%
  pivot_longer(
    cols = starts_with("x"),
    names_to = "date",
    values_to = "zhvi_price"
  ) %>%
  mutate(date = as.Date(date, format = "x%Y_%m_%d"))  
```


### Step 3: Convert ZORI Data to Long Format

```{r pivot_zori}
# Convert ZORI Data to Long Format
zori_long = nyc_data %>%
  pivot_longer(
    cols = starts_with("x"),
    names_to = "date",
    values_to = "zori_rental_price"
  ) %>%
  mutate(date = as.Date(date, format = "x%Y_%m_%d")) 
```


## Problem 2 – Quality Control and EDA

### Question 1: Explian the reason

### Question 2: Compare the number

```{r}
# Count ZIP Codes in NYC Rental Dataset
num_zip_rental = nyc_rental %>%
  distinct(zip_code) %>%
  nrow()

# Count ZIP Codes in ZIP Code Dataset
num_zip_data = zip_data %>%
  distinct(zip_code) %>%
  nrow()

# Print the counts
cat("Number of ZIP Codes in NYC Rental Dataset:", num_zip_rental, "\n")
cat("Number of ZIP Codes in ZIP Code Dataset:", num_zip_data, "\n")
```



### Question 3: Average Rental Price by Borough and Year

```{r avg_rent_by_borough}
# Calculate Average Rental Price by Borough and Year
avg_rent_by_borough_year = zori_long %>%
  mutate(year = lubridate::year(date)) %>%
  group_by(borough, year) %>%
  summarise(avg_rental_price = mean(zori_rental_price, na.rm = TRUE)) %>%
  arrange(borough, year)

# Display Table
avg_rent_by_borough_year %>%
  knitr::kable(digits = 2, caption = "Average Rental Price by Borough and Year")
```

### Question 4: Rental Price Drop During COVID-19

```{r rental_price_drop}
# Compare Rental Prices in January 2021 to January 2020
covid_rent_drop = zori_long %>%
  filter(date == as.Date("2020-01-31") | date == as.Date("2021-01-31")) %>%
  pivot_wider(names_from = date, values_from = zori_rental_price) %>%
  mutate(price_drop = `2020-01-31` - `2021-01-31`) %>%
  group_by(borough) %>%
  filter(price_drop == max(price_drop, na.rm = TRUE)) %>%
  select(borough, neighborhood, price_drop)

# Display Table
covid_rent_drop %>%
  knitr::kable(digits = 2, caption = "Largest Rental Price Drop from 2020 to 2021 by Borough")
```

## Problem 3 – Visualization

### Qusetion 1: NYC Rental Prices Over Time

```{r rental_price_plot}
# Plot NYC rental prices for each ZIP code over time
zori_long %>%
  ggplot(aes(x = date, y = zori_rental_price, color = borough)) +
  geom_line(alpha = 0.5) +
  labs(
    title = "NYC Rental Prices by ZIP Code Over Time",
    x = "Date",
    y = "Rental Price Index (ZORI)"
  ) +
  theme_minimal()
```


### Qusetion 2:2023 ZIP Code Housing Price Distribution Across States
```{r zhvi_distribution_plot}
# Calculate average housing price for each ZIP code in 2023
avg_zhvi_by_zip = zhvi_long %>%
  group_by(zip_code) %>%
  summarise(avg_zhvi_price = mean(zhvi_price, na.rm = TRUE))

# Plot the distribution of housing prices to compare across states
avg_zhvi_by_zip %>%
  ggplot(aes(x = avg_zhvi_price)) +
  geom_histogram(binwidth = 50000, fill = "skyblue", color = "black", alpha = 0.7) +
  labs(
    title = "Distribution of ZIP Code-Level Housing Prices Across States in 2023",
    x = "Average Housing Price (ZHVI)",
    y = "Number of ZIP Codes"
  ) +
  theme_minimal()
```

### Qusetion 3: 2023 NYC Rental Prices vs. Housing Prices

```{r rental_vs_housing_plot}
# Calculate the average monthly rental price and housing price for each ZIP code in 2023
zori_zhvi_2023 = zori_long %>%
  filter(lubridate::year(date) == 2023) %>%
  group_by(zip_code) %>%
  summarise(avg_rental_price = mean(zori_rental_price, na.rm = TRUE)) %>%
  left_join(
    zhvi_long %>%
      filter(lubridate::year(date) == 2023) %>%
      group_by(zip_code) %>%
      summarise(avg_zhvi_price = mean(zhvi_price, na.rm = TRUE)),
    by = "zip_code"
  )

# Plot rental prices against housing prices
zori_zhvi_2023 %>%
  ggplot(aes(x = avg_rental_price, y = avg_zhvi_price)) +
  geom_point(alpha = 0.6) +
  labs(
    title = "Comparison of Rental and Housing Prices by ZIP Code in NYC for 2023",
    x = "Average Rental Price (ZORI)",
    y = "Average Housing Price (ZHVI)"
  ) +
  theme_minimal()
```
